Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue

### 2021-05-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
pairwise group DGE with intercept

## Figure 3

>>>>

## Figure 3
## individual group comparison DGE
```{r}
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
TRX_counts <- TRX_counts[,sample_use]

metadata <- metadata %>%
  #select(Contraception, sexwork_months) %>%
  mutate(sexwork_months=ifelse(is.na(.$sexwork_months),
                               mean(metadata$sexwork_months,na.rm = TRUE),.$sexwork_months))%>%
  mutate(Contraception=ifelse(.$Contraception=="no HC", "no_HC","DMPA")) %>%
  select(ID, PatID, HIVstatus, Contraception, age, sexwork_months, everything())

#design <- model.matrix(~  group_annotation, data=metadata[sample_use,])
#design <- model.matrix(~0 + group_annotation+Contraception+HIVstatus+age+sexwork_months, 
                       #data=metadata[sample_use,])
design <- model.matrix(~group_annotation+HIVstatus+Contraception+age+sexwork_months, 
                       data=metadata[sample_use,])
colnames(design) <- sub("group_annotation","",colnames(design))

#lrt <- glmLRT(fit,coef=2:5)
lrt <- glmLRT(fit,coef=1:9)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("group_annotation","",colnames(top))
#top <- cbind(logFC.Gard=0,top)

## individual group comparision ##
# without intercept:
n <- combn(sort(unique(group_annotation)),2) 
g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
s <- seq_along(1:ncol(n)) %>% set_names(g)
#name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))

# with intercept:
n <- combn(sort(unique(group_annotation))[-1],2) 
g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
g <- c(sort(unique(group_annotation))[-1], g)

CONTRASTS <- makeContrasts( contrasts=g,
                            levels = design )
lrt <- map(s, ~ glmLRT(fit, contrast = CONTRASTS[,.x]))
top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]])

#saveRDS(top, file = "./TopTable_all_groups_06.04.2021.RDS")
#saveRDS(top, file = "./TopTable_ModifiedM4_pairwise_17.05.2021.RDS")
## end of individual group comparision ##

# save TopTables 
library(openxlsx)
l_top <- list(ModifiedM4 = top)
imap(l_top, ~write.xlsx(.x, file=paste0("../../../results/","TopTable",.y,".xlsx"))) # Excel

## individual GO comparison ##
source("../enrichment_function.R")
gmt_list_KEGG <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
gmt_list_GO <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")
enrich_list <- function(x, gmt_list){
  temp <- compute_enrichment(genes = x,
                            gmt_list = gmt_list,
                            min_terms_pathway = 10,
                            max_terms_pathway = 300,
                            min_overlap = 3,
                            sort_by_pvalue = T)
  temp <- temp[!grepl("REGULATION",rownames(temp)),]
  return(temp)
}

info_logFC <- map(top, ~table(abs(.x$logFC) > 1))
info_PValue <- map(top, ~table(.$PValue < 0.01))

# up and down regulated genes seperatly
t_split <- map(top, ~ .x %>%
        dplyr::filter(., .$PValue < 0.01) %>%
        #dplyr::filter(., abs(.$logFC) < 1) %>%
        mutate(., reg = ifelse(.$logFC > 0, "up", "down")) %>%
        split(., .$reg)
  )
# up and down regulated genes toghether
t <- map(top, ~ .x %>%
        dplyr::filter(., .$PValue < 0.01) #%>%
        #dplyr::filter(., abs(.$logFC) < 1) 
  )

## GO enrichment
enrichment_groups <- map_depth(t, 1, ~enrich_list(rownames(.x), gmt_list_GO))
enrichment_groups_split <- map_depth(t_split, 2, ~enrich_list(rownames(.x), gmt_list_GO)) %>%
  map(., ~bind_rows(.x, .id = "regulation"))

## save files
l_enrich <- list(ModifiedM4_split = enrichment_groups_split, ModifiedM4 = enrichment_groups)
imap(l_enrich, ~write.xlsx(.x, file=paste0("../../../results/","GO_enrichment_",.y,".xlsx"))) # Excel
## end of individual GO comparison ##
```
